---
title: "WFBZ Summaries"
format: html
---

---
title: "WFBZ Summaries"
format: html
execute:
  echo: false
  warning: false
  message: false
---


```{r packages}
library(sf)
library(tidyverse)
library(gt)
library(glue)
library(khroma)
library(tigris)
source("../code/helpers.R")
```


```{r data}
wfbz <- read_sf("../wflite.geojson") 

regions <- read_sf(unzip_url("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.AdministrativeRegion.zip", dst = tempfile(fileext = ".shp"))) %>%
  select(usfs_region = REGIONNAME) %>%
  mutate(usfs_region = str_replace(usfs_region, " Region", ""))
```


```{r}
# Find Nearest Region
wfbz_region <- regions$usfs_region[st_nearest_feature(wfbz, regions, pairwise = TRUE)]

wfbz_crit <- wfbz %>%
  st_drop_geometry() %>%
  select(all_of(c("wildfire_year", "wildfire_states", "wildfire_max_civil_fatalities", "wildfire_struct_destroyed", "wildfire_fema_dec", "wildfire_community_intersect", "wildfire_disaster_criteria_met"))) %>%
  mutate(usfs_region = wfbz_region) %>%
  filter(wildfire_community_intersect) %>%
  group_by(usfs_region, wildfire_year) %>%
  summarize(
  	`Events` = n(),
  	`Fatalities` = sum(wildfire_max_civil_fatalities, na.rm = TRUE),
  	`Structures Destroyed` = sum(wildfire_struct_destroyed, na.rm = TRUE),
  	`FEMA Declaration` = sum(wildfire_fema_dec, na.rm = TRUE)
  )

wfbz_crit_all <- wfbz %>%
  st_drop_geometry() %>%
  select(all_of(c("wildfire_year", "wildfire_states", "wildfire_max_civil_fatalities", "wildfire_struct_destroyed", "wildfire_fema_dec", "wildfire_community_intersect", "wildfire_disaster_criteria_met"))) %>%
  mutate(usfs_region = wfbz_region) %>%
  group_by(usfs_region, wildfire_year) %>%
  summarize(
  	`Events` = n(),
  	`Fatalities` = sum(wildfire_max_civil_fatalities, na.rm = TRUE),
  	`Structures Destroyed` = sum(wildfire_struct_destroyed, na.rm = TRUE),
  	`FEMA Declaration` = sum(wildfire_fema_dec, na.rm = TRUE)
  )


```


```{r, fig.height = 5, fig.width = 12}

plotme <- function(dat, var, rel = FALSE){
	# Determine order to plot in
	region_summary <- dat %>%
    ungroup() %>% 
    group_by(usfs_region) %>%
    summarize(tot = sum(get(var), na.rm = TRUE), .groups = 'drop') %>%
    mutate(pcttot = tot/sum(tot)) %>%
    arrange(desc(pcttot)) 
	
	region_order <- as.character(region_summary$usfs_region)
	
	# Add details to labels
	if(rel){
		region_label <- paste0(as.character(region_summary$usfs_region), ' (', format(100*region_summary$pcttot, digits = 1), '%)')
	}else{
		region_label <- paste0(as.character(region_summary$usfs_region), ' (', str_replace(format(region_summary$tot, big.mark = ','), '\\s+', ''), ')')
	}
	
	# Assign colors to regions consistently
	region_colors <- khroma::color('light', reverse = TRUE)(9)
	names(region_colors) <- sort(region_label)
	
	dat$usfs_region <- factor(dat$usfs_region, levels = region_order, labels = region_label)

	# Summarize and Plot
	if(rel){
		dat %>% 
			ggplot() +
		  geom_col(
		    aes(x = wildfire_year, y = get(var), fill = usfs_region),
		    color = "grey25",
		    size = .1,
		    width = 1,
		    position = position_fill()
		  ) +
		  scale_fill_manual(
		  	name = "USFS Region (Overall Share)",
		  	values = region_colors
		  ) +
		  scale_x_continuous(expand = expansion(mult = c(0, 0)), limits = c(1999.5, 2024.5), breaks = seq(2000, 2024, by = 2)) +
		  scale_y_continuous(expand = expansion(mult = c(0, 0)), labels = scales::percent) +
		  xlab("Year") +
		  ylab(glue("{var} (Share of Total)")) +
			#ggtitle(var) + 
		  gguw::theme_uw() +
		  theme(
		  	panel.border = element_rect(color = "grey25", fill = NA, size = .25),
		  	plot.title = element_text(size = 14),
				axis.text.x = element_text(angle = 90, vjust = +.25),
		  	axis.ticks = element_line()		  
			)
	}else{
		dat %>% 
			ggplot() +
		  geom_col(
		    aes(x = wildfire_year, y = get(var), fill = usfs_region),
		    color = "grey25",
		    size = .1,
		    width = 1
		  ) +
		  scale_fill_manual(
		  	name = "USFS Region (Overall Total)",
		  	values = region_colors
		  ) +
		  scale_x_continuous(expand = expansion(mult = c(0, 0)), limits = c(1999.5, 2024.5), breaks = seq(2000, 2024, by = 2)) +
		  scale_y_continuous(expand = expansion(mult = c(0, 0))) +
		  xlab("Year") +
		  ylab(glue("{var}")) +
			#ggtitle(var) + 
		  gguw::theme_uw() +
		  theme(
		  	panel.border = element_rect(color = "grey25", fill = NA, size = .25),
		  	plot.title = element_text(size = 14),
		  	axis.text.x = element_text(angle = 90, vjust = +.25),
		  	axis.ticks = element_line()
		  )
	}
}
```

### Number of Fire Disasters per USFS Region over Time

#### Community-Intersecting

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit, 'Events', rel = FALSE)
plotme(wfbz_crit, 'Events', rel = FALSE)
```

#### All

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit_all, 'Events', rel = FALSE)
plotme(wfbz_crit_all, 'Events', rel = TRUE)
```

### Number of Fatalities per USFS Region over Time

The five deadliest fires in our data set are labeled. 


#### Community-Intersecting

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit, 'Fatalities', rel = FALSE)  + 
	annotate("text", x = 2023, y = 50, size = 2, angle = -45, hjust = 'right', label = "Lahaina Fire (HI) ———") + 
	annotate("text", x = 2018, y = 85, size = 2, angle = -45, hjust = 'right', label = "Camp Fire (CA) ———") +
	annotate("text", x = 2017, y = 50, size = 2, angle = -45, hjust = 'right', label = "Tubbs Fire (CA) ———") +
	annotate("text", x = 2013, y = 19, size = 2, angle = -45, hjust = 'right', label = "Yarnell Hill Fire (AZ) ———") + 
	annotate("text", x = 2020, y = 20, size = 2, angle = -45, hjust = 'right', label = "North Complex Fire (CA) ———")
plotme(wfbz_crit, 'Fatalities', rel = TRUE)
```

#### All 

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit_all, 'Fatalities', rel = FALSE) + 
	annotate("text", x = 2023, y = 50, size = 2, angle = -45, hjust = 'right', label = "Lahaina Fire (HI) ———") + 
	annotate("text", x = 2018, y = 85, size = 2, angle = -45, hjust = 'right', label = "Camp Fire (CA) ———") +
	annotate("text", x = 2017, y = 50, size = 2, angle = -45, hjust = 'right', label = "Tubbs Fire (CA) ———") +
	annotate("text", x = 2013, y = 19, size = 2, angle = -45, hjust = 'right', label = "Yarnell Hill Fire (AZ) ———") + 
	annotate("text", x = 2020, y = 20, size = 2, angle = -45, hjust = 'right', label = "North Complex Fire (CA) ———")
plotme(wfbz_crit_all, 'Fatalities', rel = TRUE)
```


### Number of Structures Destroyed per USFS Region over Time

The five fires in our data set with the most destroyed structures are labeled. 

Note that the 

#### Community-Intersecting 

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit, 'Structures Destroyed', rel = FALSE) + 
	annotate("text", x = 2018, y =150*100, size = 2, angle = -45, hjust = 'right', label = "Camp Fire (CA) ———") + 
	annotate("text", x = 2017, y = 85*100, size = 2, angle = -45, hjust = 'right', label = "Tubbs Fire (CA) ———") +
	annotate("text", x = 2023, y = 35*100, size = 2, angle = -45, hjust = 'right', label = "Lahaina Fire (HI) ———") +
	annotate("text", x = 2015, y = 45*100, size = 2, angle = -45, hjust = 'right', label = "Valley Fire (CA) ———") + 
	annotate("text", x = 2018, y =140*100, size = 2, angle = -45, hjust = 'right', label = "Woolsey (CA) ———")
plotme(wfbz_crit, 'Structures Destroyed', rel = TRUE)
```


#### All 

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit_all, 'Structures Destroyed', rel = FALSE) + 
	annotate("text", x = 2018, y =150*100, size = 2, angle = -45, hjust = 'right', label = "Camp Fire (CA) ———") + 
	annotate("text", x = 2017, y = 85*100, size = 2, angle = -45, hjust = 'right', label = "Tubbs Fire (CA) ———") +
	annotate("text", x = 2023, y = 35*100, size = 2, angle = -45, hjust = 'right', label = "Lahaina Fire (HI) ———") +
	annotate("text", x = 2015, y = 45*100, size = 2, angle = -45, hjust = 'right', label = "Valley Fire (CA) ———") + 
	annotate("text", x = 2018, y =140*100, size = 2, angle = -45, hjust = 'right', label = "Woolsey (CA) ———")
plotme(wfbz_crit_all, 'Structures Destroyed', rel = TRUE)
```


### Number of FEMA Declarations per USFS Region over Time

#### Community-Intersecting 

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit, 'FEMA Declaration', rel = FALSE)
plotme(wfbz_crit, 'FEMA Declaration', rel = TRUE)
```

#### All

```{r, fig.height = 4, fig.width = 8}
plotme(wfbz_crit_all, 'FEMA Declaration', rel = FALSE)
plotme(wfbz_crit_all, 'FEMA Declaration', rel = TRUE)
```


### USFS Regions Mapped

```{r, fig.height = 4, fig.width = 8}
regions_cart <- regions %>%
	st_cast('POLYGON') %>%
	tigris::shift_geometry()

region_colors <- khroma::color('light', reverse = TRUE)(9)
names(region_colors) <- sort(unique(regions$usfs_region))

ggplot(regions_cart) + 
	geom_sf(aes(fill = usfs_region), color = NA) +  
	geom_sf(data = tigris::shift_geometry(tigris::states(cb = TRUE, progress_bar = FALSE)), color = 'black', fill = NA, size = .1) +
	scale_fill_manual(
  	name = "USFS Region",
  	values = region_colors
	) +
	scale_x_continuous(limits = c(-3115585, 2263736)) + 
	scale_y_continuous(limits = c(-1702303, 1559103)) +
	gguw::theme_uw() +
	theme(
		axis.text = element_blank()
	)
```

